Ice-shelf ocean boundary layer dynamics from large-eddy simulations

### 2021-07-14 main.tex

\documentclass[draft]{agujournal2019}
\usepackage{url} 
\usepackage{lineno}
\usepackage{graphicx}
%\usepackage{xr}
\usepackage{float}
%\externaldocument{A_suppl_fig}
\usepackage{amsmath}
\linenumbers
%\draftfalse
\journalname{JGR: Oceans}

@@

The turbulence closure scheme is Anisotropic Minimum Dissipation \cite<AMD;>{rozema_minimum-dissipation_2015} employing the extension to scalars introduced by \citeA{abkar_minimum-dissipation_2016}. We validated our implementation against the stable atmospheric boundary layer test case published in \citeA{abkar_large-eddy_2017}. Typical sub-grid scale momentum and scalar diffusivities in our sub-ice-shelf simulations of stratified turbulence range from 10\textsuperscript{-5} to 10\textsuperscript{-4} m\textsuperscript{2} s\textsuperscript{-1} in the upper two-thirds of the domain where damping is not applied (the damping methodology is discussed in Section \ref{meth:setup}.

@@

% Description of time evolution
Melt rates decline over the course of the simulations (Figure \ref{fig:timeseries}), preventing the identification of steady-state melt rates for most runs. This decline is in response to the concomitant increase in stratification during the course of the simulations, which decreases vertical heat fluxes by reducing vertical turbulent fluctuations. We do not continue our simulations beyond 4 inertial periods in the hope of reaching steady-state conditions because the increase in stratification reduces the turbulent length scales and necessitates higher model resolution than we have used here. Our simulations generally show resolved turbulent fluxes exceeding subgrid turbulent fluxes, but subgrid turbulent fluxes dominate within several meters of the ice base where the stratification is strongest (Figure S1).

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/tke_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/melt_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/us_cmp_dT_t.png}
    \end{minipage}
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/tke_cmp_dslope_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/melt_cmp_dslope_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/us_cmp_dslope_t.png}
    \end{minipage}
    \caption{Time evolution of (a,d) domain-averaged turbulence kinetic energy (TKE), (b,e) melt rate and (c,f) friction velocity for (a-c) thermal driving simulations and (d-f) variable slope simulations. The black curve represents the same simulation in all panels.}
    % Consider shading the analysis window
    % TODO add panel labels
    \label{fig:timeseries}
\end{figure}

@@

% Description of velocity profiles
Ekman rotation near the boundary can be seen in all simulations (Figs. \ref{fig:dT_profiles}c and \ref{fig:dslope_profiles}c), but for more strongly sloped runs, buoyancy plays an increasingly strong role in driving the mean flow near the boundary. This effect can be seen most clearly by comparing the up-slope component of the flow within several meters of the ice-ocean interface across the slope-varying simulations (Figure \ref{fig:dslope_profiles}c). These mean buoyancy effects increase flow velocities on the order of a few cm s\textsuperscript{-1} as slope varies from 0.01$^{\circ}$ to 1$^{\circ}$ (Figure \ref{fig:dslope_profiles}c). Flow velocities near the boundary increase on the order of 10 cm s\textsuperscript{-1} as far-field thermal driving increases from 0.5 to 0.6$^{\circ}$C at 1$^{\circ}$ slope (Figure \ref{fig:dT_profiles}c). This velocity increase is attributed to changes in the magnitude of buoyancy forcing at the boundary, related to differences in melt rates. The time-averaged far-field velocity shown in Figures~\ref{fig:dT_profiles}c and \ref{fig:dslope_profiles}c integrates inertial oscillations of the geostrophic velocity at 20 cm s\textsuperscript{-1}.

% BL depth
Stratification within the boundary layer increases with thermal driving (Figure~\ref{fig:dT_profiles}a,b). Thus, the increase in melt-induced buoyancy fluxes with thermal driving dominates over the increase in shear induced by the buoyant flow. Conversely, stratification decreases with increasing slope, indicating that the increase in shear induced by the buoyant flow dominates over the increase in melt-induced buoyancy fluxes with slope. The simulated boundary layer depth has a strong relationship with local thermal driving (i.e., the difference between IOBL temperature and the freezing point), reflecting entrainment of warm water below and setting mixing lengthscales  within the IOBL. Boundary layer depth is identified with the local maximum in the ratio of vertical to horizontal velocity variance (e.g., Figure~S1). This definition of the IOBL can also be considered the depth to which shear originating at the boundary influences the flow. In none of the simulations is the boundary layer well-mixed with respect to either temperature or salinity. The boundary layer depth increases through time, reaching 15 -- 25 m at the end of the simulations.

% TKE budget, shear
Shear production of turbulence dominates the TKE budget. The evolution of TKE can be described by three source terms and dissipation:
\begin{equation}
    \frac{d(\textrm{TKE})}{dt} = F_{shear} + F_{buoy} + F_{trans} - \textrm{diss}
\end{equation}
Figure \ref{fig:tke_budget}b-d shows the source terms in this budget for the variable slope simulations; the analogous figure for the thermal driving simulations is Figure S2 and shows similar patterns. Shear production ($F_{shear}$) ranges from $\sim 10^{-9} \textrm{--} 10^{-8}$ m\textsuperscript{2} s\textsuperscript{-3} with a maximum at the ice-ocean interface and local maxima in the upslope flow within 5 m of the boundary (Figure \ref{fig:tke_budget}a). The increase in TKE throughout the boundary layer as slope increases reflects this shear-induced turbulence (Figure \ref{fig:tke_budget}a).

@@

% TKE budget, dissipation, tendency
Dissipation ($diss$), the remaining term in the TKE budget, can be inferred from the remainder of these terms and the rate of change of TKE. In these simulations TKE is not in steady-state, with an average  dissipation rate over the course of a simulation of $\mathcal{O}(10^{-11})$ m\textsuperscript{2} s\textsuperscript{-3} as seen in Figure \ref{fig:timeseries}a,c. Dissipation in the IOBL is on the same order as shear production ($10^{-9}$ m\textsuperscript{2} s\textsuperscript{-3}), as all other terms in the TKE budget are small.

@@

\begin{figure}[]
    \centering
    \begin{minipage}{0.45\textwidth}
        \includegraphics[trim={1cm 0 3cm 0},clip,width=\textwidth]{Figures/pt_slope01_xz_y64_zmax42_t40.png}
    \end{minipage}%
    \begin{minipage}{0.55\textwidth}
        \includegraphics[trim={1.5cm 0 0.5cm 0},clip,width=\textwidth]{Figures/pt_slope1_xz_y64_zmax42_t40.png}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \includegraphics[trim={1cm 0 3cm 0}, clip, width=\textwidth]{Figures/sa_slope01_xz_y64_zmax42_t40.png}
    \end{minipage}%
    \begin{minipage}{0.55\textwidth}
        \includegraphics[trim={1.5cm 0 0.5cm 0}, clip, width=\textwidth]{Figures/sa_slope1_xz_y64_zmax42_t40.png}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \includegraphics[trim={1.5cm 0 3.4cm 0},clip,width=\textwidth]{Figures/v_slope01_xy_z1_zmax1_t40.png}
    \end{minipage}%
    \begin{minipage}{0.6\textwidth}
        \centering
        \includegraphics[trim={2cm 0 0.5cm 0},clip,width=\textwidth]{Figures/v_slope1_xy_z1_zmax1_t40.png}
    \end{minipage}%
    \begin{minipage}{0.05\textwidth}
    \hfill
    \end{minipage}
    \caption{Instantaneous flow structures observed at 40h in (a,c,e) 0.01$^{\circ}$ slope case and (b,d,f) 1$^{\circ}$ slope case. (a,b) Temperature in cross section mid-way through the y-axis. (c,d) Salinity in cross-section mid-way through the y-axis. (e,f) Cross-slope velocity fluctuations at 1 m below the ice-ocean interface. Note that the colorbar scale is smaller in (e) than (f).}
    % Consider contours to a-d to show structures more clearly
    % TODO add panel labels
    \label{fig:cross-section}
\end{figure}

@@

% intermittent turbulence
Melt rate fluctuations increase in amplitude as thermal driving increases and as slope increases, which we attribute to the increasing importance of buoyancy forcing in driving a near-boundary plume and thus determining the depth-distribution of shear. At higher slopes ($\geq 0.5^{\circ}$), inertial oscillations enhance shear mixing first near the boundary; this mixing front then propagates to the base of the mixed layer over a few hours (Figure \ref{fig:intermittency}b). The two simulations associated with the highest thermal driving values, and the highest near-boundary stratification, experience a dramatic reduction in vertical velocity fluctuations during the down-slope flow period. This can be seen in the time series of friction velocity (i.e., shear stress) at the ice interface (Figure \ref{fig:timeseries}b). Below the mixed layer, there is periodic intermittency in turbulence interrupted by the penetration of shear-induced turbulence when the IOBL buoyantly accelerates. For this reason, we do not examine IOBL entrainment in this study. We discuss these rotational effects on turbulent intensity further in Section \ref{disc:dyn}.

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.4\textwidth}
        \includegraphics[trim={0 0cm 3cm 0cm},clip,width=\textwidth]{Figures/dslope2/e_hovmoller.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 0cm 0 0cm},clip,width=\textwidth]{Figures/dT4/e_hovmoller.png}
    \end{minipage}
    \caption{Example of turbulent intermittency as seen in white regions of turbulence kinetic energy (TKE). (a) A non-intermittent case: 0.1$^{\circ}C$ thermal driving, 0.01$^{\circ}$ slope. (b) An intermittent case: 0.6$^{\circ}C$ thermal driving, 1$^{\circ}$ slope.}
    \label{fig:intermittency}
    % TODO add panel labels
\end{figure}

% Melt rate sensitivity to thermal driving
Time-averaged melt rates depend fairly linearly on far-field thermal driving (Figure \ref{fig:melt_sensitivity}a). This is mostly attributable to a linear relationship between melt rate and interfacial thermal driving, as differences in friction velocity and the thermal exchange coefficient are small and do not have a systematic relationship with far-field thermal driving. 
The derived, time-averaged thermal exchange coefficient ($\Gamma_{T,der}$) on the other hand does have a weakly negative relationship with far-field thermal driving, reflecting decrease efficiency of heat exchange with increasing near-interface stratification. We discuss these derived thermal exchange coefficients further in Section \ref{disc:prm_gamma}.

@@

% Melt sensitivity to slope
There is also a linear relationship between melt rate and ice-shelf basal slope, with threshold-like behavior at slopes less than 0.01$^{\circ}$. This linear relationship is due primarily to a linear relationship between friction velocity and melt rate while there is a smaller ($\sim\!1/3$ size) opposite effect from decreasing interfacial thermal driving with increasing slope (not shown). This threshold behavior in melting in the two lowest-slope cases is attributable to similar friction velocities (Figure \ref{fig:melt_sensitivity}d) arising from similar IOBL velocities and turbulence  (Figure \ref{fig:dslope_profiles}c,d). This behavior is discussed in more detail in Section \ref{disc:prm_sens}. The differences in derived thermal exchange coefficient are not pronounced as slope varies, on the order of a 10\% change as slope increases from 0.01$^{\circ}$ to 1$^{\circ}$ (Figure \ref{fig:melt_sensitivity}d). These differences in friction velocity arise from higher IOBL velocities and turbulence at higher slopes.

@@

% Scaling for flux profiles
The relatively narrow range of conditions simulated here suggests that a depth-dependent shape function for scalar fluxes could be formulated. The distance from the interface is scaled by the Ekman depth:
\begin{equation}
    d_E = (2 K_b)^{1/2} |f_3|^{-1/2}
\end{equation}
where $K_{b}$ is the mean bulk viscosity in the turbulent boundary layer assuming the total fluxes follow Fick's law:
\begin{equation}
    \overline{w'\mathbf{u}'} = K_{b} d\mathbf{u}/dz
\end{equation}
$K_b$ profiles are shown in Figure S6.

@@

% Scaling heat flux profiles with slope
Despite melt rates scaling reasonably well with the basal slope ($\sin \alpha$), the vertical heat flux profile does not, showing a much lower sensitivity to slope ($(\sin\alpha)^{1/4}$, Figure \ref{fig:heatflux_profiles}d). There is strong agreement between the scaled vertical heat flux profiles at different slopes, though the threshold behavior at low slopes noted in melt rates is replicated here. We also find that the Ekman depth is a poor predictor of boundary layer depth. This may be due to the depth-variable shear induced by buoyancy which is not reflected in the depth-mean IOBL bulk viscosity used to compute the Ekman depth. Our study was not designed to evaluate IOBL entrainment, which would be better explored by varying the level of background stratification, so we devote no further attention to it here. We discuss scaling of vertical fluxes and the possibility of their parameterization in Section \ref{disc:prm_vmix}.

% momentum flux profiles
By looking at the variation of momentum flux with depth shown in Figure \ref{fig:momflux_profiles}, we can examine the relative contributions of interfacial drag and buoyant acceleration of the IOBL to the velocity profiles shown in Figure \ref{fig:dT_profiles}c and \ref{fig:dslope_profiles}c. Since the mean velocity is oriented both upslope (+$u$) and cross-slope (+$v$), positive $\overline{u'w'}$ and $\overline{v'w'}$ close to the interface indicate slowing of the current due to drag whereas negative $\overline{u'w'}$ and $\overline{v'w'}$ a few m below the interface through much of the boundary layer indicate buoyant acceleration of the IOBL. The amplitude of momentum fluxes increases with slope and thermal driving consistent with increases in IOBL velocity and TKE. Intermittency in turbulence also changes the shape of momentum flux profile, displacing momentum peaks closer to the surface as the IOBL shoals (pink curve, Figure \ref{fig:momflux_profiles}a).

@@

Excess dissipation could arise either through the sub-grid scheme or the model numerics. We found a rapid loss of turbulence in PALM simulations when a dynamic Smagorinky turbulence closure was used, which is consistent with previous studies on the limited applicability of the standard Smagorinsky turbulence closure to strongly stratified flows due to the strong anisotropy of those flows \cite{flores_analysis_2011, jimenez_large-eddy_2005}. This motivated our adoption of the AMD turbulence closure scheme \cite{abkar_minimum-dissipation_2016}. However, we found that the buoyancy term added to this scheme by \citeA{abkar_large-eddy_2017} had an unrealistically high magnitude in the vicinity of the ice base where gradients are large. We removed this term as \citeA{vreugdenhil_stratification_2019} found it to be negligible, but our unrealistic solution for this term suggests that AMD scheme may not perform optimally at the resolution employed here, which is significantly lower than that used by \citeA{vreugdenhil_stratification_2019}.

@@

The flux Richardson number, $\text{Ri}_f$, the ratio of buoyancy flux to shear-driven TKE production, provides a measure of mixing efficiency. The simulated $\text{Ri}_f$ values of 0.05 -- 0.1 are well below the critical $\text{Ri}_f$ of $\sim\!0.2$, indicating that we are in the regime in which mixing efficiency ($\text{Ri}_f$) decreases with increasing stratification ($\text{Ri}_g$) \cite{armenio_investigation_2002, peltier_mixing_2003}. This is consistent with our finding that the thermal exchange coefficient decreases for the high thermal driving cases which also achieve stronger IOBL stratification.

Our simulations are certainly missing some sources of TKE present in ice-shelf cavities which could modify IOBL structure and mixing efficiency. These simulations did not include tides, which provide perturbations to the mean velocity that enhance melt rates and entrainment, especially at Filchner-Ronne \cite{makinson_modeling_1999, makinson_influence_2011, mueller_tidal_2018}. While internal gravity waves arise in LES, they may be of smaller amplitude and play a lesser role in mixing than they would in a real ice-shelf settings due to the absence of large-scale external forcings such as tides and storms, seafloor topography, as well as possible resonance with the cavity geometry \cite{gwyther_cold_2020, robertson_tidally_2013, padman_ocean_2018, mueller_impact_2012}. Enhanced drag at the ice-shelf base could increase shear production of TKE; however under strong stratification, surface roughness elements may suppress turbulence rather than enhance it \cite{ohya_wind-tunnel_2001}.

>>>>

\documentclass[draft]{agujournal2019}
\usepackage{url} 
\usepackage{lineno}
\usepackage{graphicx}
%\usepackage{xr}
\usepackage{float}
%\externaldocument{A_suppl_fig}
%\usepackage{amsmath}
\usepackage{mathtools}
\linenumbers
\draftfalse
\journalname{JGR: Oceans}

@@

The turbulence closure scheme is Anisotropic Minimum Dissipation \cite<AMD;>{rozema_minimum-dissipation_2015} employing the extension to scalars introduced by \citeA{abkar_minimum-dissipation_2016}. We validated our implementation against the stable atmospheric boundary layer test case published in \citeA{abkar_large-eddy_2017}. Typical sub-grid scale momentum and scalar diffusivities in our sub-ice-shelf simulations of stratified turbulence range from 10\textsuperscript{-5} to 10\textsuperscript{-4} m\textsuperscript{2} s\textsuperscript{-1} in the upper two-thirds of the domain where damping is not applied (the damping methodology is discussed in Section \ref{meth:setup}).

@@

% Description of time evolution
Melt rates decline over the course of the simulations (Figure \ref{fig:timeseries}), preventing the identification of steady-state melt rates for most runs. This decline is in response to the concomitant increase in stratification during the course of the simulations, which decreases vertical heat fluxes by reducing vertical turbulent fluctuations. We do not continue our simulations beyond 4 inertial periods in the hope of reaching steady-state conditions because the increase in stratification reduces the turbulent length scales and necessitates higher model resolution than we have used here. Our simulations generally show resolved turbulent fluxes exceeding subgrid turbulent fluxes, but subgrid turbulent fluxes dominate within several meters of the ice base where the stratification is strongest (Figure S1). Resolved turbulent fluxes are also comparable to subgrid turbulent fluxes late in the simulation of low slope cases after significant TKE loss (Figure S1d).

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/tke_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/melt_cmp_dT_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/us_cmp_dT_t.png}
    \end{minipage}
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/tke_cmp_dslope_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/melt_cmp_dslope_t.png}
    \end{minipage}%
    \begin{minipage}{0.33\textwidth}
        \includegraphics[trim={0 0 0 0},clip,width=\textwidth]{Figures/us_cmp_dslope_t.png}
    \end{minipage}
    \caption{Time evolution of (a,d) domain-averaged, resolved turbulence kinetic energy (TKE), (b,e) melt rate and (c,f) friction velocity for (a-c) thermal driving simulations and (d-f) variable slope simulations. The black curve represents the same simulation in all panels. The time axis excludes the initial spin-up period. }
    % Consider shading the analysis window
    % TODO add panel labels
    \label{fig:timeseries}
\end{figure}

@@

% Description of velocity profiles
Ekman rotation near the boundary can be seen in all simulations (Figs. \ref{fig:dT_profiles}c and \ref{fig:dslope_profiles}c), but for more strongly sloped runs, buoyancy plays an increasingly strong role in driving the mean flow near the boundary. This effect can be seen most clearly by comparing the up-slope component of the flow within several meters of the ice-ocean interface across the slope-varying simulations (Figure \ref{fig:dslope_profiles}c). These mean buoyancy effects increase flow velocities on the order of a few cm s\textsuperscript{-1} as slope varies from 0.01$^{\circ}$ to 1$^{\circ}$ (Figure \ref{fig:dslope_profiles}c). Flow velocities near the boundary increase on the order of 10 cm s\textsuperscript{-1} as far-field thermal driving increases from 0.5 to 0.6$^{\circ}$C at 1$^{\circ}$ slope (Figure \ref{fig:dT_profiles}c). This velocity increase is attributed to changes in the magnitude of buoyancy forcing at the boundary, related to differences in melt rates. The time-averaged far-field velocity shown in Figures~\ref{fig:dT_profiles}c and \ref{fig:dslope_profiles}c integrates inertial oscillations of the geostrophic velocity with a characteristic magnitude of 20 cm s\textsuperscript{-1}.

% BL depth
In none of the simulations do we observe a well-mixed the boundary layer well-mixed with respect to scalars, with varying degrees of stratification over the first 20 m from the boundary. Thus, we characterize the IOBL using a definition of the mixing layer depth as depth at which the dissipation rate drops below 10\textsuperscript{-9} m\textsuperscript{2} s\textsuperscript{-3} \cite{franks_has_2014, sutherland_mixed_2014}. Since shear is the primary source of TKE as discussed later in this section, the IOBL here can also be thought of as portion of the water column where shear originating at the boundary influences the flow. The boundary layer depth increases through time, reaching 15 -- 25 m at the end of the simulations.

Stratification within the boundary layer increases with thermal driving (Figure~\ref{fig:dT_profiles}a,b). Thus, the increase in melt-induced buoyancy fluxes with thermal driving dominates over the increase in shear induced by the buoyant flow. Conversely, stratification decreases with increasing slope, indicating that the increase in shear induced by the buoyant flow dominates over the increase in melt-induced buoyancy fluxes with slope. The simulated boundary layer depth has a strong relationship with local thermal driving (i.e., the difference between IOBL temperature and the freezing point), reflecting entrainment of warm water below and setting mixing lengthscales  within the IOBL.

% TKE budget, shear
Shear production of turbulence dominates the TKE budget. The evolution of resolved TKE can be described by three source terms and dissipation:
\[
    \frac{de}{dt} = 
        \underbrace{
            -\overline{u^{\prime}w^{\prime}}\frac{du}{dz} -\overline{v^{\prime}w^{\prime}}\frac{dv}{dz}}
        _{F_{shear}} - 
        \underbrace{
            \frac{g_z}{\rho}\overline{\rho^{\prime}w^{\prime}}        -\frac{g_x}{\rho}\overline{\rho^{\prime}u^{\prime}}}
        _{F_{buoy}} + 
        \underbrace{
            \frac{d}{dz}(\overline{w^{\prime}p^{\prime}} +
                     \overline{w}^{\prime}e^{\prime})}
        _{F_{trans}} -
        \hspace{-2mm}\underbrace{\frac{}{}\vareps}_{diss}
\]

Figure \ref{fig:tke_budget}b-d shows the source terms in this budget for the variable slope simulations; the analogous figure for the thermal driving simulations is Figure S2 and shows similar patterns. Shear production ($F_{shear}$) ranges from $\sim 10^{-9} \textrm{--} 10^{-8}$ m\textsuperscript{2} s\textsuperscript{-3} with a maximum at the ice-ocean interface and local maxima in the upslope flow within 5 m of the boundary (Figure \ref{fig:tke_budget}a). The increase in TKE throughout the boundary layer as slope increases reflects this shear-induced turbulence (Figure \ref{fig:tke_budget}a).

@@

% TKE budget, dissipation, tendency
Dissipation ($diss$), the remaining term in the TKE budget, can be inferred from the remainder of these terms and the rate of change of TKE. Since we evaluate resolved TKE ($e$), dissipation here represents the transfer of energy to the sub-grid scales; there is no flow of energy from sub-grid to resolved scales. In these simulations TKE is not in steady-state, with an average  dissipation rate over the course of a simulation of $\mathcal{O}(10^{-11})$ m\textsuperscript{2} s\textsuperscript{-3} as seen in Figure \ref{fig:timeseries}a,c. Dissipation in the IOBL is on the same order as shear production ($10^{-9}$ m\textsuperscript{2} s\textsuperscript{-3}), as all other terms in the TKE budget are small.

@@

\begin{figure}[]
    \centering
    \begin{minipage}{0.45\textwidth}
        \includegraphics[trim={1cm 0 3cm 0},clip,width=\textwidth]{Figures/pt_slope01_xz_y64_zmax42_t40.png}
    \end{minipage}%
    \begin{minipage}{0.55\textwidth}
        \includegraphics[trim={1.5cm 0 0.5cm 0},clip,width=\textwidth]{Figures/pt_slope1_xz_y64_zmax42_t40.png}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \includegraphics[trim={1cm 0 3cm 0}, clip, width=\textwidth]{Figures/sa_slope01_xz_y64_zmax42_t40.png}
    \end{minipage}%
    \begin{minipage}{0.55\textwidth}
        \includegraphics[trim={1.5cm 0 0.5cm 0}, clip, width=\textwidth]{Figures/sa_slope1_xz_y64_zmax42_t40.png}
    \end{minipage}
    \begin{minipage}{0.45\textwidth}
        \includegraphics[trim={1.5cm 0 3.4cm 0},clip,width=\textwidth]{Figures/v_slope01_xy_z1_zmax1_t40.png}
    \end{minipage}%
    \begin{minipage}{0.6\textwidth}
        \centering
        \includegraphics[trim={2cm 0 0.5cm 0},clip,width=\textwidth]{Figures/v_slope1_xy_z1_zmax1_t40.png}
    \end{minipage}%
    \begin{minipage}{0.05\textwidth}
    \hfill
    \end{minipage}
    \caption{Instantaneous flow structures observed at 40h in (a,c,e) 0.01$^{\circ}$ slope case and (b,d,f) 1$^{\circ}$ slope case. (a,b) Temperature in cross section mid-way through the y-axis. (c,d) Salinity in cross-section mid-way through the y-axis. (e,f) Cross-slope velocity fluctuations at 1 m below the ice-ocean interface. Note that the colorbar scale is smaller in (e) than (f).}
    % TODO add panel labels
    \label{fig:cross-section}
\end{figure}

@@

% intermittent turbulence
Melt rate fluctuations increase in amplitude as thermal driving increases and as slope increases, which we attribute to the increasing importance of buoyancy forcing in driving a near-boundary plume and thus determining the depth-distribution of shear. At higher slopes ($\geq 0.5^{\circ}$), inertial oscillations enhance shear mixing first near the boundary; this mixing front then propagates to the base of the IOBL over a few hours (Figure \ref{fig:intermittency}b). The two simulations associated with the highest thermal driving values, and the highest near-boundary stratification, experience a dramatic reduction in vertical velocity fluctuations during the down-slope flow period. This can be seen in the time series of friction velocity (i.e., shear stress) at the ice interface (Figure \ref{fig:timeseries}b). In all simulations, there is periodic intermittency in turbulence below the IOBL interrupted by the penetration of shear-induced turbulence when the IOBL buoyantly accelerates. We discuss these rotational effects on turbulent intensity further in Section \ref{disc:dyn}.

\begin{figure}[h!]
    \centering
    \begin{minipage}{0.4\textwidth}
        \includegraphics[trim={0 0cm 3cm 0cm},clip,width=\textwidth]{Figures/dslope2/e_hovmoller.png}
    \end{minipage}%
    \begin{minipage}{0.5\textwidth}
        \includegraphics[trim={0 0cm 0 0cm},clip,width=\textwidth]{Figures/dT4/e_hovmoller.png}
    \end{minipage}
    \caption{Example of turbulent intermittency as seen in white regions of turbulence kinetic energy (TKE). (a) A non-intermittent case: 0.1$^{\circ}C$ thermal driving, 0.01$^{\circ}$ slope. (b) An intermittent case: 0.6$^{\circ}C$ thermal driving, 1$^{\circ}$ slope.}
    \label{fig:intermittency}
    % TODO add panel labels, add IOBL contour
\end{figure}

% Melt rate sensitivity to thermal driving
Time-averaged melt rates depend fairly linearly on far-field thermal driving (Figure \ref{fig:melt_sensitivity}a). This is mostly attributable to a linear relationship between melt rate and interfacial thermal driving, as differences in friction velocity and the thermal exchange coefficient are small and do not have a systematic relationship with far-field thermal driving. On the other hand, the derived, time-averaged thermal exchange coefficient representing the efficiency of heat exchange from -2 m depth to the ice base ($\Gamma_{T,der}$) does have a weakly negative relationship with far-field thermal driving (Figure \ref{fig:melt_sensitivity}c). This indicates a decrease in the efficiency of heat exchange with increasing near-interface stratification. The highest thermal driving case shows an anomalously high thermal exchange coefficient over the last inertial period, which features high TKE shown in Figure \ref{fig:intermittency}b. We discuss these derived thermal exchange coefficients further in Section \ref{disc:prm_gamma}.

@@

% Melt sensitivity to slope
There is also a linear relationship between melt rate and ice-shelf basal slope, with threshold-like behavior at slopes less than 0.01$^{\circ}$. This linear relationship is due primarily to a linear relationship between friction velocity and melt rate while there is a smaller ($\sim\!1/3$ size) opposite effect from decreasing interfacial thermal driving with increasing slope (not shown). These differences in friction velocity arise from higher IOBL velocities and turbulence at higher slopes. We observe threshold behavior in melting in the two lowest-slope cases, can be attributed to similar friction velocities arising from similar IOBL velocities and turbulence (Figure \ref{fig:melt_sensitivity}d and Figure \ref{fig:dslope_profiles}c,d). This behavior is discussed in more detail in Section \ref{disc:prm_sens}. The differences in derived thermal exchange coefficient are not pronounced as slope varies, on the order of a 10\% change as slope increases from 0.01$^{\circ}$ to 1$^{\circ}$ (Figure \ref{fig:melt_sensitivity}d).

@@

% Scaling for flux profiles
The relatively narrow range of conditions simulated here suggests that a depth-dependent shape function for scalar fluxes could be formulated. The distance from the interface is scaled by the Ekman depth:
\begin{equation}
    d_E = (2 K_e)^{1/2} |f_3|^{-1/2}
\end{equation}
where $K_e$ is the mean eddy viscosity in the turbulent boundary layer assuming the total fluxes follow Fick's law:
\begin{equation}
    \overline{w'\mathbf{u}'} = K_e d\mathbf{u}/dz
\end{equation}
$K_e$ profiles are shown in Figure~S6.

@@

% Scaling heat flux profiles with slope
Despite melt rates scaling reasonably well with the basal slope ($\sin \alpha$), the vertical heat flux profile does not, showing a much lower sensitivity to slope ($(\sin\alpha)^{1/4}$, Figure \ref{fig:heatflux_profiles}d). There is strong agreement between the scaled vertical heat flux profiles at different slopes, though the threshold behavior at low slopes noted in melt rates is replicated here. We also find that the Ekman depth is a poor predictor of boundary layer depth. This may be due to the depth-variable shear induced by buoyancy which is not reflected in the depth-mean IOBL eddy viscosity used to compute the Ekman depth. Our study was not designed to evaluate IOBL entrainment, which would be better explored by varying the level of background stratification, so we devote no further attention to it here. We discuss scaling of vertical fluxes and the possibility of their parameterization in Section \ref{disc:prm_vmix}.

% momentum flux profiles
By looking at the variation of momentum flux with depth shown in Figure \ref{fig:momflux_profiles}, we can examine the relative contributions of interfacial drag and buoyant acceleration of the IOBL to the velocity profiles shown in Figure \ref{fig:dT_profiles}c and \ref{fig:dslope_profiles}c. Since the mean velocity is oriented both upslope (+$u$) and cross-slope (+$v$), positive $\overline{u'w'}$ and $\overline{v'w'}$ close to the interface indicate slowing of the current due to drag whereas negative $\overline{u'w'}$ and $\overline{v'w'}$ a few m below the interface through much of the boundary layer indicate buoyant acceleration of the IOBL. The amplitude of momentum fluxes increases with slope and thermal driving consistent with increases in IOBL velocity and TKE.

@@

Excess dissipation could arise either through the sub-grid scheme or the model numerics. We found a rapid loss of turbulence in PALM simulations when a dynamic Smagorinky turbulence closure was used, which is consistent with previous studies on the limited applicability of the standard Smagorinsky turbulence closure to strongly stratified flows due to the strong anisotropy of those flows \cite{flores_analysis_2011, jimenez_large-eddy_2005}. This motivated our adoption of the AMD turbulence closure scheme \cite{abkar_minimum-dissipation_2016}. However, we found that the buoyancy term added to this scheme by \citeA{abkar_large-eddy_2017} had an unrealistically high magnitude in the vicinity of the ice base where gradients are large. We removed this term as \citeA{vreugdenhil_stratification_2019} found it to be negligible, but our unrealistic solution for this term suggests that AMD scheme may not perform optimally at the resolution employed here, which is significantly lower than that used by \citeA{vreugdenhil_stratification_2019} but enables us to extend the domain allow for the development of a $\sim$20 m-thick IOBL.

@@

The flux Richardson number, $\text{Ri}_f$, the ratio of buoyancy flux to shear-driven TKE production, provides a measure of mixing efficiency. The simulated $\text{Ri}_f$ values of 0.05 -- 0.1 are well below the critical $\text{Ri}_f$ of $\sim\!0.25$, indicating that we are in the regime in which mixing efficiency ($\text{Ri}_f$) decreases with increasing stratification ($\text{Ri}_g$) \cite{armenio_investigation_2002, peltier_mixing_2003}. This is consistent with our finding that the derived, 2 m-depth thermal exchange coefficient decreases for the high thermal driving cases which also achieve stronger IOBL stratification.

Our simulations are certainly missing some sources of TKE present in ice-shelf cavities which could modify IOBL structure and mixing efficiency. These simulations did not include tides, which provide perturbations to the mean velocity that enhance melt rates and entrainment, especially at Filchner-Ronne \cite{makinson_modeling_1999, makinson_influence_2011, mueller_tidal_2018}. While internal gravity waves arise in LES, in our LES model they may be of smaller amplitude and play a lesser role in mixing than they would in a real ice-shelf settings due to the absence of large-scale external forcings such as tides and storms, seafloor topography, as well as possible resonance with the cavity geometry \cite{gwyther_cold_2020, robertson_tidally_2013, padman_ocean_2018, mueller_impact_2012}. Enhanced drag at the ice-shelf base could increase shear production of TKE; however under strong stratification, surface roughness elements may suppress turbulence rather than enhance it \cite{ohya_wind-tunnel_2001}.
